### 2025-08-26 chapters/introduction.tex
solo revisions complete

ARM TrustZone~\cite{suzaki2021ts, turn0search8} takes a different approach by partitioning the processor into secure and normal worlds, enabling lightweight isolation for Trusted Applications but with generally weaker isolation guarantees than SGX. Its performance impact tends to be lower, yet TrustZone’s capabilities and APIs are less standardized, which complicates broad adoption and benchmarking~\cite{turn0search9}.

AMD SEV and newer technologies such as SEV-SNP offer memory encryption and integrity protections for virtual machines, providing scalable enclave-like isolation for cloud workloads. Coppolino et al.~\cite{turn0search2_coppolino} experimentally evaluated SEV and TDX against SGX, revealing performance and transparency trade-offs that influence their suitability for different application scenarios.

In parallel, open-source efforts like Keystone~\cite{Lee2019, dayeol2019keystone} aim to democratize TEE development by providing a modular framework on RISC-V processors, supporting flexible enclave configurations and custom security policies. Research into Keystone’s memory models and sharing mechanisms~\cite{yu2022elasticlave, lee2022cerberus} has demonstrated improved performance and formally verified security properties, addressing common TEE bottlenecks such as enclave memory management overhead.

Benchmarking and performance measurement frameworks are crucial to evaluate these TEEs fairly and holistically. TS-perf~\cite{suzaki2021ts} offers a unified performance suite targeting multiple TEE architectures, including SGX, TrustZone, and Keystone, facilitating comparative studies. Other benchmarking efforts include SGXGauge~\cite{kumar2022sgxgauge}, which specifically assesses Intel SGX overheads under different workload characteristics, and traditional CPU benchmarks like Dhrystone~\cite{weiss2002dhrystone, york2002benchmarking} and CoreMark~\cite{gal2012exploring}, adapted for enclave contexts.

>>>>

ARM TrustZone~\cite{suzaki2021tsperf} takes a different approach by partitioning the processor into secure and normal worlds, enabling lightweight isolation for Trusted Applications but with generally weaker isolation guarantees than SGX. Its performance impact tends to be lower, yet TrustZone’s capabilities and APIs are less standardized, which complicates broad adoption and benchmarking~\cite{turn0search9}.

AMD SEV and newer technologies such as SEV-SNP offer memory encryption and integrity protections for virtual machines, providing scalable enclave-like isolation for cloud workloads. Coppolino et al.~\cite{coppolino2025experimental} experimentally evaluated SEV and TDX against SGX, revealing performance and transparency trade-offs that influence their suitability for different application scenarios.

In parallel, open-source efforts like Keystone~\cite{Lee2019} aim to democratize TEE development by providing a modular framework on RISC-V processors, supporting flexible enclave configurations and custom security policies. Research into Keystone’s memory models and sharing mechanisms~\cite{yu2022elasticlave, lee2022cerberus} has demonstrated improved performance and formally verified security properties, addressing common TEE bottlenecks such as enclave memory management overhead.

Benchmarking and performance measurement frameworks are crucial to evaluate these TEEs fairly and holistically. TS-perf~\cite{suzaki2021tsperf} offers a unified performance suite targeting multiple TEE architectures, including SGX, TrustZone, and Keystone, facilitating comparative studies. Other benchmarking efforts include SGXGauge~\cite{kumar2022sgxgauge}, which specifically assesses Intel SGX overheads under different workload characteristics, and traditional CPU benchmarks like Dhrystone~\cite{weiss2002dhrystone, york2002benchmarking} and CoreMark~\cite{gal2012exploring}, adapted for enclave contexts.
